A Systematic Literature Review on Role-Based Access Control

### 2013-02-19 paper/rbac-slr.tex
Some minor updates and the start of a data extraction section.
Adding terms section.
Updates to results, process.
Adds analysis section for research questions.
Adds outline for tracking. Adds some preliminary research question sections and analysis. Adds more to results section.

\documentclass[preprint,12pt]{elsarticle}

\usepackage{amssymb}

@@

%% SECTIONS %%
\input{sections/intro}
\input{sections/process}

>>>>

\documentclass[3p,12pt]{elsarticle}

\usepackage{amssymb}
\usepackage{hyperref}

@@

\section{Outline}

\begin{enumerate}
\item Introduction
    \begin{itemize}
  	\item RBAC history, RBAC as a standard
  	\item Audience that should care about our review
  	\item Brief explanation of what has led to extensions cropping up
  	\item Paper organization
    \end{itemize}
\item Background
    \begin{itemize}
   	\item What is RBAC
   	\item Core RBAC and it's entities, and what it tries to solve
    \end{itemize}
\item Process
\item Results
    \begin{itemize}
  	\item Data used in process, results of process for paper count
  	\item Categorization that came out applying process (based on word usage)
  	\item Results for each research question (raw data)
    \end{itemize}
\item Analysis
    \begin{itemize}
  	\item Explanation and definitions of categories based on results
  	\item Examination of results in relation to each research question and larger trends drawn from data
  	\item Examination of cross research question concerns an
    \end{itemize}
\item Conclusion
\end{enumerate}

%% SECTIONS %%
\input{sections/intro}
\input{sections/terms}
\input{sections/process}
\input{sections/results}
\input{sections/analysis}

### 2013-02-19 paper/sections/analysis.tex
Adds outline for tracking. Adds some preliminary research question sections and analysis. Adds more to results section.

P1: Discuss focus on going through each research question individually, and combining results across at the end of the process

P2: List the research questions

S1: RQ1 and discuss by category and general results

S2: RQ2 and discuss by category and general results

S3: RQ3 and discuss by category and general results

S4: RQ4 and discuss by category and general results

S5: RQ5 and discuss by category and general results

>>>>

\subsection{Research Question 5}

Do the extension models have corresponding implementations in practice? \\

When designing and proposing a model targeted at a feature that is rooted in practical
usage by real software systems, bringing the model to life is strong evidence that the
proposed model can work in practice.  The concept of authorization, and access control
is rooted in a business need.  Thus, any access control model needs to be feasible
in the real world not just on paper.  We analyzed the primary sources to see how many
proposed models actually had implmenetations associated with them.  And quantified the
type of implementation.  Whether the implementation was for a real system, for a prototype
and/or used in a production environment.

Of the 29 papers surveyed, there was a significant lack of implementation with 21 of the paper providing no
mention of a implementation or prototype.  Of the remianing 8 papers that did mention an implementation, half 
were simply prototypes developed by the authors while the other half were claimed to be implemented within a real
system.

\subsection{Research Question 6}

How are extended RBAC models evaluated in theory and in practice? \\

Providing evaluation of a proposed model is a key component in establishing the models validity.  The papers were examined
for evidence of evaluations ranging from performance to mathematical accuracy to application to real world scenarios.  Further,
for each proposed model, the reviewers looked for evidence that the authors made comparisons between their own model and the base
model as they pertained to claims made by the authors of why their model is needed.  The quantifiable evaluations looked for were:

\begin{itemize}
\item Time-based Performance
\item Complexity analysis
\item Comparison to standard RBAC
\item Mathematical accuracy
\item Example scenarios of the model in action
\item Experimental analysis of the model
\item Case study of the model in practice
\end{itemize}

Based on the diverse evaluation criteria, 12 models presented no evidence of an evaluation.  8 models presented example scenarios
and how application of their model would apply and resolve the situation.  6 of the models provided some form of performance
or complexity analysis of their model.  This included graphs of the model's time to determine authorization as the number of entities
grew, and the size of the role space for the extension model compared to standard RBAC. 4 models provided mathematical descriptions
and analsysis as a way to provide evaluation in the form of completeness.

The most widely used evaluation method was providing sample scenarios with accompanying workflows of how the extension model
would tackle those scenarios.  Much is left to the reader to assume of these types of evaluations, as the authors do not explicitly state
or show how the standard model is deficient in tackling said scenarios.

\subsection{Research Question X}

What domains or scenarios serve as inspiration for these extensions? \\

Business needs have historically driven RBAC research and development.  The primary mode of evaluation for
model extensions has been the presentation of business scenarios in various domains and how the model
uniquely handles those particular scenarios.  Thus, looking for trends in the domains used in the example
scenarios might serve to illuminate a trend worth further examination into the reason for the explosion of
RBAC extensions.

Upon examination of the primary sources, the most prevalent domains were:

\begin{itemize}
\item What domains or scenarios serve as inspiration for these extensions?
\item Medical domain
\item Pervasive computing environments
\item Mobile devices
\item Large-scale organizations with many sub-departments
\item Enterprise, organization workflows
\end{itemize}

### 2013-02-19 paper/sections/intro.tex
Adds outline for tracking. Adds some preliminary research question sections and analysis. Adds more to results section.

Since the introduction of the standardized RBAC model, innovation and the spread of software engineering into new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle.  For example, RBAC is ill-equipped to handle the additional entities that need to be taken into consideration during authorization with the introduction of privacy concerns in domains such as healthcare  Thus, a series of extensions to the standard RBAC model have appeared in the literature, each adding one or more features.

Since standard RBAC model has limitations such as specifying environmental constraints or context information
Researchers developed extended models of RBAC to overcome the limitations.
However, as researchers often develop their own specialized extended models of RBAC,
their research cannot be generalized or compared with other research work appropriately.
As a result, researchers could take time on reinventing the wheel.
But how do we, as a community, ensure that a metric is suitable and acceptable for its intended purpose?

The goal of this work is to provide practioners an assessment of the state of extended models of RBAC..

We conducted a systematic literature review (SLR) to evaluate and interpret all available research relevant to a particular research question or topic area of interest.

Our research provides contributions to the community by:

>>>>

Since the introduction of the standardized RBAC model, innovation and the spread of software engineering into new domains has led to scenarios that, by some accounts, the standard RBAC model cannot handle.  For example, RBAC is ill-equipped to handle the additional entities that need to be taken into consideration during authorization with the introduction of privacy concerns in domains such as healthcare  Thus, a series of extensions to the standard RBAC model have appeared in the literature, each adding one or more features.  However, as researchers often develop their own specialized extended models of RBAC, their research cannot be generalized or compared with other research work appropriately.

A systematic literature review is designed to yield quanitfiable results across a body of research in order to draw out similarities, trends and defiencies in an area as a whole.  The review we have performed should serve as a future starting place for researcher's looking to tackle new problems in the realm of authorization and prevent re-invention of the wheel. For developer's looking to find a model to fit their seemingly unique access control needs, this review will provide a basis for comparison and easy look-up for what extension based model to use.

The goal of this work is to provide practioners an assessment of the state of extended models of RBAC\ldots

Our research provides contributions to the community in the following ways:
